### 2018-09-28 rendering-templates/word_template.rmd
adds rendering template

---
title: "Rmd to word template here are some words to make this title run into two pages and words"
author: "Gaurav Kandlikar"
date: "April 18, 2018"
output:  
  word_document:
    reference_docx: word_template.docx
bibliography: ../bibliography.bib
---

# Single title

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

Here's a citation to [@Bever_1997].

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

### Here's a third heading

```{r cars}
summary(cars)
```

#### Here's a fourth heading

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## References